---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# icecdr

<!-- badges: start -->
<!-- badges: end -->

The goal of icecdr is to download sea ice concentration data from the NSIDC Climate Data Record. V4. 

## Installation

You can install the development version of icecdr from [GitHub](https://github.com/) with:

``` r
# install.packages("pak")
pak::pak("eliocamp/icecdr")
```

## Example

Use the `cdr_*` functions to download satellite-derived Antarctic or Arctic sea ice concentration data at monthly or daily resolution. 

The basic usage downloads a NetCDF file in a temporary directory.

```{r example}
library(icecdr)

dates <- c("2020-01-01", "2023-01-01")
cdr_antarctic_monthly(dates, dir = "data", use_cache = TRUE)
```


With `use_cache = TRUE`, data will not be downloaded if needed. 

```{r}
system.time(cdr_antarctic_monthly(dates, dir = "data", use_cache = TRUE))
```

There are four simple functions to download whole-domain data:

```r
cdr_antarctic_daily()
cdr_antarctic_monthly()

cdr_arctic_daily()
cdr_arctic_monthly()
```

But the `cdr()` function exposes all arguments to download any crazy subset

```r
cdr(date_range = dates,
    # Data every 7 days
    date_stride = 3,
    resolution = "daily", 
    # Thin the grid by taking every other gridpoint
    xgrid_stride = 2,
    ygrid_stride = 2, 
    hemisphere = "north"
    )
```

The `cdr_fix_grid()` function will fix the grid information to make it work with [CDO](https://code.mpimet.mpg.de/projects/cdo). This requires CDO installed in your system. 

```{r}
library(rcdo)
library(ggplot2)

cdr_antarctic_monthly(dates, dir = "data", use_cache = TRUE) |> 
  cdr_fix_grid() |> 
  cdo_fldint() |> 
  cdo_execute() |> 
  metR::ReadNetCDF(c(aice = "cdr_seaice_conc_monthly")) |> 
  ggplot(aes(time, aice)) +
  geom_line() +
  labs(y = "Sea ice area")
  
```

