---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file --> 

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# icecdr

<!-- badges: start -->
[![Codecov test coverage](https://codecov.io/gh/eliocamp/icecdr/graph/badge.svg)](https://app.codecov.io/gh/eliocamp/icecdr)
[![R-CMD-check](https://github.com/eliocamp/icecdr/actions/workflows/R-CMD-check.yaml/badge.svg)](https://github.com/eliocamp/icecdr/actions/workflows/R-CMD-check.yaml)
<!-- badges: end -->

The goal of icecdr is to download sea ice concentration data from the NSIDC Climate Data Record. 

## Installation

You can install the icecdr with:

``` r
install.packages("icecdr")
```

## Example

Use the `cdr_*` functions to download satellite-derived Antarctic or Arctic sea ice concentration data at monthly or daily resolution. 
They use version 5 by default but version 4 is also supported. 

The basic usage downloads a NetCDF file in a temporary directory.

```{r example}
library(icecdr)

dates <- c("2020-01-01", "2023-01-01")
cdr_antarctic_monthly(dates, dir = "data")
```


By default, files are only downloaded if needed.

```{r}
system.time(cdr_antarctic_monthly(dates, dir = "data"))
```

There are four simple functions to download whole-domain data:

```r
cdr_antarctic_daily()
cdr_antarctic_monthly()

cdr_arctic_daily()
cdr_arctic_monthly()
```

But the `cdr()` function exposes all arguments to download any crazy subset

```r
cdr(date_range = dates,
    # Data every 7 days
    date_stride = 7,
    resolution = "daily", 
    # Thin the grid by taking every other gridpoint
    xgrid_stride = 2,
    ygrid_stride = 2, 
    hemisphere = "north"
    )
```

The `cdr_fix()` function will fix the grid information to make it work with [CDO](https://code.mpimet.mpg.de/projects/cdo) and standardise variable names. This requires CDO installed in your system. 

```{r}
library(rcdo)
library(ggplot2)

extent_cdr <- cdr_antarctic_monthly(dates, dir = "data") |> 
  cdr_fix() |> 
  cdo_gtc(0.15) |>
  cdo_fldint() |> 
  cdo_execute() |> 
  metR::ReadNetCDF(c(extent = "aice")) 

extent_index <- sea_ice_index("south", "monthly", dir = "data") |> 
  data.table::fread() |>
  subset(time >= dates[1] & time <= dates[2])

extent_cdr |>
  ggplot(aes(time, extent)) +
  geom_line(aes(colour = "cdr")) +
  geom_line(data = extent_index, aes(colour = "sea_ice_index")) +
  labs(y = "Sea ice extent")
  
```

